Data preparation - individual

### 2025-07-29 procesamiento/01-prod_prep-individual.Rmd
data for reproducibility

---
title: "Data preparation"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

```{r eval=FALSE, include=FALSE}
knitr::knit(input = here::here("procesamiento/01-prod_prep-micro.Rmd"));file.remove("01-prod_prep-micro.md")
knitr::knit(input = here::here("procesamiento/02-prod_prep-macro.Rmd"));file.remove("02-prod_prep-macro.md")
```

@@

```{r}
#frq(df1$v59)
df1$hhdepriv <- car::recode(df1$v59,recodes = "c(0,8,9)=NA")
df1$hhdepriv <- set_na(df1$hhdepriv,na = c(0,8,9))
# df1$hhdepriv <- sjmisc::rec(x = df1$hhdepriv,rec = "rev")
sjlabelled::set_label(df1$hhdepriv) <- "Household economic deprivation (household difficulties)"
#frq(df1$hhdepriv)

#frq(df1$v11)
df1$perineq <- car::recode(df1$v11,recodes = "c(0,8,9)=NA")
df1$perineq <- set_na(df1$perineq,na = c(0,8,9))
df1$perineq <- sjmisc::rec(x = df1$perineq,rec = "rev")
sjlabelled::set_label(df1$perineq) <- "Perceived economic inequality (Inc. diff. too large)"
#frq(df1$perineq)

#frq(df1$v12)
df1$prefineq <- car::recode(df1$v12,recodes = "c(0,8,9)=NA")
df1$prefineq <- set_na(df1$prefineq,na = c(0,8,9))
df1$prefineq <- sjmisc::rec(x = df1$prefineq,rec = "rev")
sjlabelled::set_label(df1$prefineq) <- "Preference inequality (society has to be fair)"
#frq(df1$prefineq)

#frq(df1$v14)
df1$pertransfer <- car::recode(df1$v14,recodes = "c(0,8,9)=NA")
df1$pertransfer <- set_na(df1$pertransfer,na = c(0,8,9))
df1$pertransfer <- sjmisc::rec(x = df1$pertransfer,rec = "rev")
sjlabelled::set_label(df1$pertransfer) <- "Perception about welfare (people become lazy)"
#frq(df1$pertransfer)
```

# Index egalitarian attitudes
```{r}
df1$egal3<- rowMeans(df1[,c("redist","prefineq","perineq")],na.rm = T)
df1$egal3 <- ((df1$egal3-min(df1$egal3,na.rm = T))/(max(df1$egal3,na.rm = T)-min(df1$egal3,na.rm = T)))*100

sjlabelled::set_label(df1$egal3) <- "Redistributive attitudes (3 items)"
# #hist.default(df1$egal3)

df1$egal2<- rowMeans(df1[,c("redist","prefineq")],na.rm = T)
df1$egal2 <- ((df1$egal2-min(df1$egal2,na.rm = T))/(max(df1$egal2,na.rm = T)-min(df1$egal2,na.rm = T)))*100
sjlabelled::set_label(df1$egal2) <- "Redistributive attitudes (2 items)"
# #hist.default(df1$egal2)
```

# Trust

```{r}
# sjmisc::frq(df1$v35)
df1$socialtrust <- car::recode(df1$v35,recodes = "c(0,8,9)=NA")
df1$socialtrust <- set_na(df1$socialtrust,na = c(0,8,9))
df1$socialtrust <- sjmisc::rec(x = df1$socialtrust,rec = "rev")
# sjmisc::frq(df1$socialtrust)
# hist.default(df1$socialtrust)
```

@@

frq(df1$is_supervisor_spo)
frq(df1$self_employed_spo)
frq(df1$n_employees_spo)

haven::write_dta(data = df1[c("id","isco88r","semp","supvis",
                              "is_supervisor","self_employed","n_employees",
                              "isco88spo","sempspo","supvisspo",
                              "is_supervisor_spo","self_employed_spo","n_employees_spo")],
                 path = here::here("input/data/proc/ocup_respondent-spouse.dta"))
```

```{r class-from-stata1}
# Load the data created in stata------------------------------------------------
df_class <- haven::read_dta(file = here::here("input/data/proc/class_respondent-spouse.dta"))
df1 <- left_join(df1,df_class,by='id')

# Class R´s and Spouse----------------------------------------------------------
frq(df1$class10)

# df1$class10 <- car::recode(df1$class10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$class10f <- as_factor(df1$class10)
frq(df1$class10f)

df1$class10f <- factor(df1$class10f,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

df1$class10f <- as_labelled(df1$class10f)
sjlabelled::set_label(df1$class10f) <- "Social Class - respondent"
frq(df1$class10f)

df1$class10spof <- as_factor(df1$class10spo)
frq(df1$class10spof)

df1$class10spof <-factor(df1$class10spof,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

df1$class10spof <- as_labelled(df1$class10spof)
sjlabelled::set_label(df1$class10spof) <- "Social Class - spouse";frq(df1$class10spof)
#if NA in class of respondent, replace with spouse

frq(df1$class10f)
frq(df1$class10spof)

df1$class11 <- ifelse(is.na(df1$class10f),df1$class10spof,df1$class10f)
frq(df1$class11)

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$class11d <- pmax(df1$class11, #social class respondent + spouse vs:
                     df1$class10spof,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$class11d)
df1$class11d <- 
  sjlabelled::set_labels(df1$class11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                         labels =  c("I", "II", "IIIab", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
sjlabelled::set_label(df1$class11d) <- "Social Class"
frq(df1$class11d)

# Class Scheme 3 classes
df1$class3 <- car::recode(df1$class11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$class3) <- "Social Class Dominance (isko - v1 - V to Intermediate)";frq(df1$class3)

df1$class3_V <- car::recode(df1$class11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3_V) <- "Social Class Dominance (isko - v1 - V to Working Class)";frq(df1$class3_V)

# Class Scheme 6 classes
df1$class6 <- car::recode(
  df1$class11d,
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

# Class only R's --------------------------------------------------------------
# Class Scheme 3 classes
df1$class3res <- car::recode(as.numeric(df1$class10f),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$class3res) <- "Social Class 3 (R's) - (isko - v1 -  V to Intermediate)";frq(df1$class3res)

df1$class3res_V <- car::recode(as.numeric(df1$class10f),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3res_V) <- "Social Class 3 (R's) - (isko - v1 -  V to Working)";frq(df1$class3res_V)

# Class Scheme 6 classes
df1$class6res <- car::recode(
  as.numeric(df1$class10f),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$class6res) <- "Social Class 6 (R's) - (isko - v1 -  V to Intermediate)";frq(df1$class6res)
# Class only Spouse --------------------------------------------------------------

# Class Scheme 3 classes
df1$class3spo <- car::recode(as.numeric(df1$class10spof),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:10='Working Class (VI+VII)'")

sjlabelled::set_label(df1$class3spo) <- "Social Class 3 (Spouse) - (isko - v1 -  V to Intermediate)";frq(df1$class3spo)

df1$class3spo_V <- car::recode(as.numeric(df1$class10spof),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:10='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$class3spo_V) <- "Social Class 3 (R's) - (isko - v1 -  V to Working)";frq(df1$class3spo_V)

# Class Scheme 6 classes
df1$class6spo <- car::recode(
  as.numeric(df1$class10spo),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)
sjlabelled::set_label(df1$class6spo) <- "Social Class 6 (Spouse) - (isko - v1)";frq(df1$class6spo) 
```

```{r class-from-stata2}
# Class R´s and Spouse----------------------------------------------------------
frq(df1$digclass10)

# df1$class10 <- car::recode(df1$class10, "7=8; 8=9; 9=10; 10=11; 11=7")
df1$digclass10f <- as_factor(df1$digclass10)
frq(df1$digclass10f)

df1$digclass10f <- factor(df1$digclass10f,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))
df1$digclass10f <- as_labelled(df1$digclass10f)
sjlabelled::set_label(df1$digclass10f) <- "Social Class - respondent"
frq(df1$digclass10f)
df1$digclass10spof <- as_factor(df1$digclass10spo)
frq(df1$digclass10spof)

df1$digclass10spof <-factor(df1$digclass10spof,
                       levels = c("higher controllers","lo controllers",
                                 "routine nonmanual",
                                 "sempl with emp","sempl without empl","selfempl farm",
                                 "manual supervisor","skilled manual","semi-unskilld manual","farm labor"),
                       labels = c("I", "II", "IIIab", "IVa", "IVb", "IVc", "V", "VI", "VIIa", "VIIb"))

df1$digclass10spof <- as_labelled(df1$digclass10spof)
sjlabelled::set_label(df1$digclass10spof) <- "Social Class - spouse"

frq(df1$digclass10spof)
#if NA in class of respondent, replace with spouse

frq(df1$digclass10f)
frq(df1$digclass10spof)

df1$digclass11 <- ifelse(is.na(df1$digclass10f),df1$digclass10spof,df1$digclass10f)
frq(df1$class11)

# Replace the highest class by dominance criteria 
# Keep the highest of two values:
df1$digclass11d <- pmax(df1$digclass11, #social class respondent + spouse vs:
                     df1$digclass10spof,#social class spouse
                     na.rm = T) # if NA, do nothing
frq(df1$digclass11d)
df1$digclass11d <- 
  sjlabelled::set_labels(df1$digclass11d,
                         levels = c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10),
                         labels =  c("I", "II", "IIIab", "IVa", "IVb", 
                                     "IVc", "V", "VI", "VIIa", "VIIb")) 
sjlabelled::set_label(df1$digclass11d) <- "Social Class 11 (isko - v2)"
frq(df1$digclass11d)

# Class Scheme 3 classes
df1$digclass3 <- car::recode(df1$digclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$digclass3) <- "Social Class 3 (isko - v2 - V to Intermediate) ";frq(df1$digclass3)

df1$digclass3_V <- car::recode(df1$digclass11d,as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3_V) <- "Social Class 3 (V to Working Class - isko - v2)";frq(df1$digclass3_V)

# Class only R's --------------------------------------------------------------
# Class Scheme 3 classes
df1$digclass3res <- car::recode(as.numeric(df1$digclass10f),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$digclass3res) <- "Social Class 3 (R's) - (V to Intermediate - isko - v2)";frq(df1$digclass3res)

df1$digclass3res_V <- car::recode(as.numeric(df1$digclass10f),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3res_V) <- "Social Class 3 (V to Working Class - isko - v2)";frq(df1$digclass3res_V)

# Class Scheme 6 classes
df1$digclass6res <- car::recode(
  as.numeric(df1$digclass10),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)

sjlabelled::set_label(df1$digclass6res) <- "Social Class 6 (R's) - (isko - v2)"

# Class only Spouse --------------------------------------------------------------
# Class Scheme 3 classes
df1$digclass3spo <- car::recode(as.numeric(df1$digclass10spo),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV+V)',
                                      'Working Class (VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:7='Intermediate class (III+IV+V)';
                             8:11='Working Class (VI+VII)'")

sjlabelled::set_label(df1$digclass3spo) <- "Social Class 3 (Spouse) - (V to Intermediate - isko - v2)";frq(df1$digclass3spo)

df1$digclass3spo_V <- car::recode(as.numeric(df1$digclass10spof),as.factor = T,
                            levels =c('Service Class (I+II)',
                                      'Intermediate class (III+IV)',
                                      'Working Class (V+VI+VII)'), 
                            "1:2='Service Class (I+II)';
                             3:6='Intermediate class (III+IV)';
                             7:11='Working Class (V+VI+VII)'")

sjlabelled::set_label(df1$digclass3spo_V) <- "Social Class 3 (Spouse) (V to Working Class - isko - v2)";frq(df1$digclass3spo_V)

# Class Scheme 6 classes
df1$digclass6spo <- car::recode(
  as.numeric(df1$digclass10spo),
  as.factor = T,
  levels = c(
    'Service class (I - higher grade)',
    'Service class (II - lower grade)',
    'Routine non-manual (IIIa+b)',
    'Petty-bourgeoise (IVa+b+c)',
    'Skilled manual workers and supv. (V+VI)',
    'Non-skilled manual workers (VIIa+b)'
  ),
  "1='Service class (I - higher grade)';
   2='Service class (II - lower grade)';
   3:4='Routine non-manual (IIIa+b)';
   5:7='Petty-bourgeoise (IVa+b+c)';
   8:9='Skilled manual workers and supv. (V+VI)';
   10:11='Non-skilled manual workers (VIIa+b)'"
)
sjlabelled::set_label(df1$digclass6spo) <- "Social Class  6 (Spouse) - (isko - v2)"
```

# Class using R

```{r digclass-egp}
# install.packages("devtools") 
# devtools::install_git("https://code.europa.eu/digclass/digclass.git")

library("DIGCLASS")
library(dplyr)

df1 <- 
df1 %>% 
mutate(egp11 = isco88_to_egp(isco88r, self_employed, n_employees, label = F,to_factor = T),
       egp11spo = isco88_to_egp(isco88spo, self_employed_spo, n_employees_spo, label = F,to_factor = T))

@@

# number of high status contacts
df1$n_high <- 
  rowSums(x = select(df1,c("lawyer","hrmanager","ceo")),na.rm = T)
sjlabelled::set_label(df1$n_high) <- 'Number of high status contacts'
# sjPlot::plot_#frq(df1$n_high,type = 'density')

# percentage of high status contacts

df1$pct_high <- (df1$n_high/df1$know_total)*100
sjlabelled::set_label(df1$df1$pct_high ) <- 'Percentage of high status contacts'
# sjPlot::plot_#frq(df1$pct_high,type = 'density')

# number of middle status contacts
df1$n_middle <-
  rowSums(x = select(df1,c("teacher","police","nurse")),na.rm = T)
sjlabelled::set_label(df1$n_middle) <- 'Number of middle status contacts'
# sjPlot::plot_#frq(df1$n_middle,type = 'density')

# percentage of middle status contacts
df1$pct_middle <- (df1$n_middle/df1$know_total)*100
sjlabelled::set_label(df1$pct_middle) <- 'Percentage of middle status contacts'
# sjPlot::plot_#frq(df1$pct_middle,type = 'density')

# number of low status contacts
# df1$n_low <- 
#   rowSums(x = select(df1,c("n_waiter","n_carmech","n_taxi","n_stvend","n_cleaner")),na.rm = T)
df1$n_low <- 
  rowSums(x = select(df1,c("mechanic","busdriver","hairbarber","cleaner")),na.rm = T)
sjlabelled::set_label(df1$n_low) <- 'Number of low status contacts'
# sjPlot::plot_#frq(df1$n_low,type = 'density')

# percentage of low status contacts
df1$pct_low <- (df1$n_low/df1$know_total)*100
sjlabelled::set_label(df1$pct_low) <- 'Percentage of low status contacts'
# sjPlot::plot_#frq(df1$pct_low,type = 'density')

# Average ISEI 
df1$isei_avg <- rowMeans(x = select(df1,isei_ocupnames),na.rm = T) #REVISAR NO DEBERIA CONTENER 0
sjlabelled::set_label(df1$isei_avg) <- 'Average contact prestige'
summary(df1$isei_avg)

# sjPlot::plot_#frq(df1$isei_avg,type = 'density')

# Total ISEI 
df1$isei_tot <- rowSums(x = select(df1,isei_ocupnames),na.rm = T)
sjlabelled::set_label(df1$isei_tot) <- 'Total contact prestige'
summary(df1$isei_tot)
# sjPlot::plot_#frq(df1$isei_tot,type = 'density')

# skimr::skim(df1$isei_tot/df1$n_total)
# skimr::skim(df1$isei_tot/df1$know_total)
# skimr::skim(df1$isei_avg)

# max ISEI 
df1 <- df1 %>% 
mutate(isei_max = do.call(pmax, c(select(., isei_ocupnames), na.rm = TRUE)))
sjlabelled::set_label(df1$isei_max) <- 'Maximum contact prestige'
# sjmisc::frq(df1$isei_max)
# sjPlot::plot_#frq(df1$isei_max)

# min ISEI
df1 <- df1 %>% 
mutate(isei_min = do.call(pmin, c(select(., isei_ocupnames), na.rm = TRUE)))
sjlabelled::set_label(df1$isei_min) <- 'Minimum contact prestige'
# sjmisc::frq(df1$isei_min)
# sjPlot::plot_#frq(df1$isei_min)

# range ISEI
table(df1$know_total>=2)
df1$isei_range <- ifelse(df1$know_total>=2,(df1$isei_max - df1$isei_min),no = NA)

sjlabelled::set_label(df1$isei_range) <- 'Range contact prestige'
# sjmisc::frq(df1$isei_range)
# sjPlot::plot_#frq(df1$isei_range)
# df1 %>% select(starts_with("isei_")) %>%View

cor(df1[,c("isei_avg","isei_max","isei_range",
                 "know_total","n_high","n_middle","n_low",
                 "pct_high","pct_middle","pct_low")],use = "complete.obs")
```

```{r heterogeneity, eval=FALSE, include=FALSE}
isei_ocupnames<- names(select(df1,busdriver:teacher))

df1$hindex_sta<- 
as.numeric(
  1 - (
    #p1^2
    (df1$n_high / df1$know_total)^2 +
    #p2^2
    (df1$n_middle / df1$know_total)^2 +
    #p3^2
    (df1$n_low / df1$know_total)^2
    )
  )

sjlabelled::set_label(df1$hindex_sta) <- "D-index: status groups"

df1$hindex_10<- 
as.numeric(
  1 - (
    #p1^2
    (df1$busdriver/df1$know_total)^2 +
    #p2^2
    (df1$ceo / df1$know_total)^2 +
    #p3^2
    (df1$cleaner / df1$know_total)^2 +
    #p3^2
    (df1$hairbarber / df1$know_total)^2 +
     #p3^2
    (df1$hrmanager / df1$know_total)^2 +
    #p3^2
    (df1$lawyer / df1$know_total)^2+
    #p3^2
    (df1$mechanic / df1$know_total)^2+
    #p3^2
    (df1$nurse / df1$know_total)^2+
    #p3^2
    (df1$police / df1$know_total)^2+    
    #p3^2
    (df1$teacher / df1$know_total)^2        
    )
  )

sjlabelled::set_label(df1$hindex_10) <- "D-index: 10 ocupations"
```

@@

```{r homogeneity-v01-V-to-intermediate}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(class3,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$class3=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3=="Intermediate class (III+IV+V)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3=="Working Class (VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

df1$homclass <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass) <- "Class homogeneity (isko v1 - V to Intermediate)"
df1$homclass_gc <- group_center(x = df1$homclass,grp = df1$country)

## Using RESPONDENT Class for homogeneity---------------------------------------
# EI homophily index
df1$ingroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Service Class (I+II)",df1$n_high,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res=="Working Class (VI+VII)",df1$n_low,df1$ingroup_res)
# frq(df1$ingroup_res)

df1$outgroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV+V)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Working Class (VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup_res)
# frq(df1$outgroup_res)

df1$homclass_res <- as.numeric(df1$ingroup_res/df1$know_total)
sjlabelled::set_label(df1$homclass_res) <- "Class homogeneity (R's) (isko v1 - V to Intermediate)"
```

```{r homogeneity-v01-V-to-working}
df1$ingroup <- NA 
df1 %>% select(class3_V,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$class3_V=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1$outgroup <- ifelse(df1$class3_V=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3_V=="Intermediate class (III+IV)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$class3_V=="Working Class (V+VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

df1$homclass_V <- as.numeric(df1$ingroup/df1$know_total)
sjlabelled::set_label(df1$homclass_V) <- "Class homogeneity (isko v1 - V to Working)"

## Using RESPONDENT Class for homogeneity---------------------------------------
df1$ingroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res_V=="Service Class (I+II)",df1$n_high,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res_V=="Intermediate class (III+IV)",df1$n_middle,df1$ingroup_res)
df1$ingroup_res <- ifelse(df1$class3res_V=="Working Class (V+VI+VII)",df1$n_low,df1$ingroup_res)
# frq(df1$ingroup_res)

df1$outgroup_res <- NA 
df1 %>% select(class3res,n_high,n_middle,n_low,outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Intermediate class (III+IV+V)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup_res)
df1$outgroup_res <- ifelse(df1$class3res=="Working Class (VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup_res)
# frq(df1$outgroup_res)

df1$homclass_V_res <- as.numeric(df1$ingroup_res/df1$know_total)
sjlabelled::set_label(df1$homclass_V_res) <- "Class homogeneity (R's) (isko v1 - V to Working)"
```

- Using my edited original coding. I recoded the variables from n_employees y self_employed. The class scheme changed a bit, where the Service class was overestimated. This should be better:

```{r homogeneity-v02-V-intermediate}
# EI homophily index
df1$ingroup <- NA 
df1 %>% select(digclass3,n_high,n_middle,n_low,ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Service Class (I+II)",df1$n_high,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Intermediate class (III+IV+V)",df1$n_middle,df1$ingroup)
df1$ingroup <- ifelse(df1$digclass3=="Working Class (VI+VII)",df1$n_low,df1$ingroup)
#frq(df1$ingroup)

df1$outgroup <- NA 
df1 %>% select(digclass3,n_high,n_middle,n_low,outgroup)
df1$outgroup <- ifelse(df1$digclass3=="Service Class (I+II)",rowSums(df1[,c("n_low","n_middle")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$digclass3=="Intermediate class (III+IV+V)",rowSums(df1[,c("n_low","n_high")],na.rm = T),df1$outgroup)
df1$outgroup <- ifelse(df1$digclass3=="Working Class (VI+VII)",rowSums(df1[,c("n_middle","n_high")],na.rm=T),df1$outgroup)
#frq(df1$outgroup)

>>>>

---
title: "Data preparation - individual"
date: "`r format(Sys.time(), '%A %d %B %Y %X')`"
output: 
  html_document: 
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: no
      number_sections: yes
    code_folding: show  
    number_sections: yes
editor_options: 
  chunk_output_type: console
---

@@

```{r}
df1$redist <- car::recode(df1$v13,recodes = "c(0,8,9)=NA")
df1$redist <- set_na(df1$redist,na = c(0,8,9))
df1$redist <- sjmisc::rec(x = df1$redist,rec = "rev")
sjlabelled::set_label(df1$redist) <- "Preferences goverment redistribution"

df1$prefineq <- car::recode(df1$v12,recodes = "c(0,8,9)=NA")
df1$prefineq <- set_na(df1$prefineq,na = c(0,8,9))
df1$prefineq <- sjmisc::rec(x = df1$prefineq,rec = "rev")
sjlabelled::set_label(df1$prefineq) <- "Preference inequality (society has to be fair)"

df1$pertransfer <- car::recode(df1$v14,recodes = "c(0,8,9)=NA")
df1$pertransfer <- set_na(df1$pertransfer,na = c(0,8,9))
df1$pertransfer <- sjmisc::rec(x = df1$pertransfer,rec = "rev")
sjlabelled::set_label(df1$pertransfer) <- "Perception about welfare (people become lazy)"
```

# Index egalitarian attitudes
```{r}
df1$egal2<- rowMeans(df1[,c("redist","prefineq")],na.rm = T)
df1$egal2 <- ((df1$egal2-min(df1$egal2,na.rm = T))/(max(df1$egal2,na.rm = T)-min(df1$egal2,na.rm = T)))*100
sjlabelled::set_label(df1$egal2) <- "Redistributive attitudes (2 items)"
```

@@

frq(df1$is_supervisor_spo)
frq(df1$self_employed_spo)
frq(df1$n_employees_spo)
```

# Social Class

```{r digclass-egp}
# install.packages("devtools") 
# devtools::install_git("https://code.europa.eu/digclass/digclass.git")
library("DIGCLASS")
df1 <- 
df1 %>% 
mutate(egp11 = isco88_to_egp(isco88r, self_employed, n_employees, label = F,to_factor = T),
       egp11spo = isco88_to_egp(isco88spo, self_employed_spo, n_employees_spo, label = F,to_factor = T))

@@

# number of high status contacts
df1$n_high <- 
  rowSums(x = select(df1,c("lawyer","hrmanager","ceo")),na.rm = T)
sjlabelled::set_label(df1$n_high) <- 'Number of high status contacts'

# number of middle status contacts
df1$n_middle <-
  rowSums(x = select(df1,c("teacher","police","nurse")),na.rm = T)
sjlabelled::set_label(df1$n_middle) <- 'Number of middle status contacts'

# number of low status contacts
df1$n_low <- 
  rowSums(x = select(df1,c("mechanic","busdriver","hairbarber","cleaner")),na.rm = T)
sjlabelled::set_label(df1$n_low) <- 'Number of low status contacts'

# Average ISEI 
#REVIEW SHOULD NOT CONTAIN 0
df1$isei_avg <- rowMeans(x = select(df1,isei_ocupnames),na.rm = T) 
sjlabelled::set_label(df1$isei_avg) <- 'Average contact prestige'
summary(df1$isei_avg)
```

@@
